---
title: "Final Project"
author: "Ri Liu (rl4508)"
date: "1/20/2022"
output: pdf_document
---


# 1. Introduction

Covid-19 has been raging for more than 2 years, causing millions of people dead, billions of people infected and incalculable economic loss. 

In these 2 years, there are several peaks of case surges caused by different variants of virus. I will analyze the trend of Covid-19 in the US from January 2020 to January 2022, including exploring the data in formats of daily new cases & deaths, 7-day average new cases & deaths and accumulated incidence & mortality rate. Plotting is a good way to show a long term trend and the difference across different states. 

Public health intervention is a complex system, except for medicine and epidemiology research, it is also affected by politics, economics and publicity. I will test whether political leaning influence pandemic development across different states. The 2020 presidential election happened right amid the pandemic. So, it is a good opportunity to research the connection of epidemic and politics. 



# 2. Packages

```{r load_packages, message=FALSE, warning=FALSE}
library(tidyverse) # data manipulation, tidying and visualization
library(cowplot) # organize plots in grid
library(car) # Levene's test for equal variance assumption
```

# 3. Data

## 3.1 Data Importing

- Covid-19 Data in the United States <https://github.com/nytimes/covid-19-data> 
- US County Level Election Results <https://github.com/tonmcg/US_County_Level_Election_Results_08-20>
- US Population Data from GPH-GU 2183
- US Standard State Name Data from R `datasets`

```{r message=FALSE}
us <- read_csv('us.csv')
states <- read_csv('us-states.csv')
counties <- read_csv('us-counties.csv')
population <- read_rds('us_state_populations.rds') 
election <- read_csv('2020_US_County_Level_Presidential_Results.csv')
```

## 3.2 Data Cleaning

- `states` and `counties`

In the `states` and `counties` data set, there are `r length(unique(states$state))` states or territorial governments. But not all of them have other information that is easy to find. So I only use the states that are in the  `state.name`  dataset and the  'District of Columbia'. 
```{r}
all_states <- append(state.name, 'District of Columbia')
#all_states
```

The districts that are temporarily ignored are as follows. This also verifies that other states' names in these data sets are in the same format. 
```{r}
except <- unique(counties$state)[!unique(counties$state) %in% all_states]
except
```

```{r}
states <- states %>%
  inner_join(tibble(state = all_states), by = 'state') %>%
  arrange(date, state)
#head(states, n = 5)
```

```{r}
counties <- counties %>%
  inner_join(tibble(state = all_states), by = 'state') %>%
  arrange(date, state, county)
#head(counties, n = 5)
```
Now, `states` and `counties` both have `r length(unique(states$state))` states. 

- Missing value

Missing value occurs in `county` column of `counties` data set when health departments report the patient's county of residence as unknown or pend determination. The editor mark these data as 'Unknown'. For the purpose of county-wise analysis, I filter out those entries with 'Unknown' `county`. Before the filtering, `counties` has `r nrow(counties)` rows of entries. 

```{r}
counties <- counties %>%
  filter(county != 'Unknown')
```

Now, `counties` has `r nrow(counties)` rows of entries.

To check whether a data set contains standard `NA`. 
```{r eval=FALSE}
check_na <- function(x) {
  any(is.na(x))
}
map(list(us, states, counties), apply, MARGIN = 2, FUN = check_na)
```
There are `NA`s in the `fips` column. `fips` is a code of standard geographic identifier for states and counties. 

```{r}
unique(counties[is.na(counties$fips), c('county', 'state')])
```
The `NA`s for 'New York City' occurred because the editor combines New York, Kings, Queens, Bronx and Richmond into 'New York City'. 'Kansas City' and 'Joplin' are separated from their origin counties. However those three new "counties" don't have their `fips` in the FIPS database. So I drop the `fips` column. 

```{r}
counties <- counties %>%
  select(!fips)
```



- `population`

In the `population` data set, there are population of each states from `r min(population$year)` to `r max(population$year)`. Only the most recent numbers are needed. 

```{r}
population <- population[population$year == max(population$year), c('state', 'population')]
```

Check the consistency of states' names in `all_states` and `population`. 
```{r}
population$state[!unique(population$state) %in% all_states]
```
There is a tiny discrepancy in the spelling of 'District of Columbia'. 
So, I need to unify the format of 'District of Columbia'. 
```{r}
population$state <- fct_recode(population$state, 'District of Columbia' = 'District Of Columbia')
```

Check again. 
```{r}
length(!unique(population$state) %in% all_states)
```



- Data Summary

Now `us` looks like:
```{r}
head(us, 5)
```

`states` looks like:
```{r}
head(states, 5)
```

`counties` looks like:
```{r}
head(counties,5)
```

`population` looks like:
```{r}
head(population, 5)
```


```{r eval=FALSE}
map(list(us = us, states = states, counties = counties), summary)
```
The `us`, `states` and `conties` now have data from `r min(us$date)` to `r max(us$date)`. 

`states` reaches its highest cases to `r max(states$cases)` and highest deaths to `r max(states$deaths)`. `counties` reaches its high cases to `r max(counties$cases)` and highest deaths to `r max(counties$deaths)`.

`election` looks like:
```{r}
head(election, 5)
```
`election` has records of `r length(unique(election$state_name))` different states, and `r length(unique(election$county_name))` different counties across the country. It records the votes count, votes difference and votes percent of each party in each county.  


# 4. Exploratory Data Analysis

## 4.1 Feature Engineering

1. Transform accumulated cases and deaths into daily new cases and new deaths. 

For `us`:
```{r}
new_case <- c(us$cases[1]) # result vector for case
new_death <- c(us$deaths[1]) # result vector for death
for (i in 2:nrow(us)){
  new_case = append(new_case, us$cases[i] - us$cases[i-1])
  new_death = append(new_death, us$deaths[i] - us$deaths[i-1])
}
us <- add_column(us, new_case, new_death)
```

For `states`:
```{r}
states <- add_column(states, new_case = c(0)) # result vector for case
states <- add_column(states, new_death = c(0)) # result vector for death
for (state in all_states){
  index = states$state == state
  states$new_case[index][1] = states$cases[index][1]
  states$new_death[index][1] = states$deaths[index][1]
  for (i in 2:nrow(states[index, ])){
    states$new_case[index][i] = states$cases[index][i] - states$cases[index][i-1]
    states$new_death[index][i] = states$deaths[index][i] - states$deaths[index][i-1]
  }
}
```


2. 7-day Average 

As each day's test ability is not all on the same level, for example, affected by holidays. Calculating 7-day average number can reduce the fluctuation to some extent. 

For `us`:
```{r}
avg7_case <- c(us$cases[1:6]) # result vector for case
avg7_death <- c(us$deaths[1:6]) # result vector for death
for (i in 7:nrow(us)){
  left <- i-6 # the first day of the seven days in a row
  avg7_case = append(avg7_case, mean(us$new_case[left:i]))
  avg7_death = append(avg7_death, mean(us$new_death[left:i]))
}
us <- add_column(us, avg7_case, avg7_death)
```
For `states`:
```{r}
states <- add_column(states, avg7_case = c(0)) # result vector for case
states <- add_column(states, avg7_death = c(0)) # result vector for death
for (state in all_states){
  index = states$state == state # filter out the state
  states$avg7_case[index][1:6] = states$new_case[index][1:6]
  states$avg7_death[index][1:6] = states$new_death[index][1:6]
  for (i in 7:nrow(states[index, ])){
    left <- i-6 # the first day of the seven days in a row
    states$avg7_case[index][i] = mean(states$new_case[index][left:i])
    states$avg7_death[index][i] = mean(states$new_death[index][left:i])
  }
}
```

3. Incidence and Mortality Rate

Transform accumulated cases and deaths into incidence and mortality rate given corresponding population. 

For `us`:
```{r}
total_population <- sum(population$population)
us <- us %>%
  mutate(incidence = cases / total_population * 1e6, 
         mortality = deaths / total_population * 1e6)
```

For `states`
```{r}
states <- states %>%
  inner_join(population, by = 'state') %>%
  mutate(incidence = cases / population * 1e6, 
         mortality = deaths / population * 1e6) 
```


Out of `states`, it is not easy to read all states' information at the same time. It is more efficient to analyse a part of the data that is more representative, which is pulling out the states with top 5 and bottom 5 incidence. However, incidence is easily affected in states with smaller population. So finally filtering out states with larger than 25% quantile population, which is `r quantile(population$population, 0.25)`, is necessary for analysis on incidence. 

```{r}
top5 <- states %>%
  filter(date == as.Date('2022-01-11'), population > 1.7*1e6) %>%
  arrange(desc(incidence)) %>%
  .[1:5, 'state', drop=T]
top5
```

```{r}
bottom5 <- states %>%
  filter(date == as.Date('2022-01-11'), population > 1.7e6) %>%
  arrange(incidence) %>%
  .[1:5, 'state', drop=T]
bottom5
```



Now, `us` looks like 
```{r}
head(us, 5)
```

`states` looks like
```{r}
head(states, 5)
```

4. 'Election' Data

The `election` data is collected on county-level. I need to summarize it to state-level, and mark the final vote for party of each state as a factor. 

```{r}
election <- election %>%
  group_by(state_name) %>%
  summarise(sum_gop = sum(votes_gop), sum_dem = sum(votes_dem)) %>%
  mutate(per_gop = sum_gop / (sum_gop+sum_dem), per_dem = sum_dem / (sum_gop+sum_dem)) %>%
  mutate(diff = round((per_gop - per_dem) * 100, 2)) %>%
  mutate(party = factor(ifelse(diff > 0, 'gop', 'dem'))) %>%
  mutate(diff = abs(diff))
```

Now, `election` looks like:
```{r}
head(election, 5)
```
If the vote difference is not significantly different, the political leaning in that state may not have a uniform characteristic. For my analysis, I filter out the states whose vote difference is larger than 25% quantile of all of the vote difference across the country, and pull out the states who vote for Republican as `gop` and who vote for Democratic as `dem`.

```{r}
quantile(election$diff, seq(0,0.5,0.05))
```

```{r}
election <- election %>%
  select(state_name, diff, party) %>%
  filter(diff > 8.650)
gop <- election %>%
  filter(diff > 8.650, party == 'gop') %>%
  .$state_name
dem <- election %>%
  filter(diff > 8.650, party == 'dem') %>%
  .$state_name
```



## 4.2 Trending Analysis

Now, I can examine the trend of COVID-19 on country-level. The scales of cases and deaths are not on the same level, so I multiply deaths by 50 to make a more readible plot.

```{r}
us %>%
  ggplot() + 
  geom_line(aes(x = date, y = avg7_case, color = 'Cases')) +
  geom_line(aes(x = date, y = avg7_death*50, color = 'Deaths * 50')) + 
  # customize the legend
  scale_color_manual(values = c('Cases' = 'blue', 'Deaths * 50' = 'red')) + 
  ylab('7-Day Average Number') + 
  ggtitle('Figure 1: 7-Day Average Number across the Country')
```
We can see that from January 2020 to January 2022, there are three major peaks in case number, which are in 2020's winter, 2021's fall and 2021's winter. And there are three major peaks in death number, two of them come with the surges of cases, but the first peak occurs at the early stage when COVID-19 was discovered, which is more severe than the case number. The death surges are mostly later than the case surges, which is easy to understand. 

However, out of the outbreak in 2021's winter, even though the case number break through an exceedingly high level compared with previous surges, the death number in this outbreak does not show a corresponding volume of surge. It could be because that the variant virus in this outbreak is less lethal than previous variants. But it could also be because that it is too early to see a surge of death in the development of this outbreak. 

Now, let's look at the trend on state-level. As the scale of the outbreak in 2021' winter is too large, I split the plot into two part. 

```{r}
# before 2021-10-31
g1 <- ggplot(filter(states, date<as.Date('2021-10-31')), aes(x = date, y = avg7_case, group = state)) + 
  geom_line(size = 0.1, alpha = 0.1) +
  geom_line(data = filter(states, state %in% top5, date<as.Date('2021-10-31')), 
            aes(date, avg7_case, group = state, color = 'Top5'), 
            color = 'red', size = 0.3, alpha = 0.5) + 
  geom_line(data = filter(states, state %in% bottom5, date<as.Date('2021-10-31')), 
            aes(date, avg7_case, group = state, color = 'Bottom5'), 
            color = 'blue', size = 0.3, alpha = 0.5) + 
  xlab('Jan 2020 - Oct 2021') + 
  ylab('7-day Average Cases') + 
  # adjust the margin between the plots
  theme(aspect.ratio = 1, plot.margin = unit(c(0,0,0,0), 'cm')) 

# after 2021-10-31
g2 <- ggplot(filter(states, date>as.Date('2021-10-31')), aes(x = date, y = avg7_case, group = state, color = 'Others')) + 
  geom_line(size = 0.1, alpha = 0.1) +
  geom_line(data = filter(states, state %in% top5, date>as.Date('2021-10-31')), 
            aes(date, avg7_case, group = state, color = 'Top5'), 
            size = 0.3, alpha = 0.5) + 
  geom_line(data = filter(states, state %in% bottom5, date>as.Date('2021-10-31')), 
            aes(date, avg7_case, group = state, color = 'Bottom5'), 
            size = 0.3, alpha = 0.5) + 
  xlab('Nov 2021 - Jan 2022') +
  ylab(label = '') + 
  theme(axis.text = element_text(size = 5)) + 
  theme(aspect.ratio = 2, plot.margin = unit(c(0,0,0,0), 'cm')) +
  # customize the legend
  scale_color_manual(values = c('Top5' = 'red', 'Bottom5' = 'blue', 'Others' = 'black')) 

# make an overall title
title <- ggdraw() + draw_label("Figure 2: 7-Day Average Cases by States", fontface='bold')

# organize the two plot in a grid 
p <- plot_grid(plot_grid(g1), plot_grid(g2), rel_widths = c(0.55, 0.45))
plot_grid(title, p, ncol = 1, rel_heights = c(0.1,1))
```
In these two plot, we can see that the states that have top 5 incidence do have higher 7-day average case number in the left plot, which is between January 2020 to October 2021. However, out of the left plot, we cannot say the same thing. The states that have better defense to the pandemic, relatively speaking, don't have significantly different performance in the outbreak of 2021's winter. 



## 4.3 Political Analysis
To research whether states with different leaning towards parties have different epidemiology results, I make a T-test. 

First, pull out the most recent incidence data and join the presidential election vote result on it. 

```{r}
plt_states <- states %>%
   filter(date == as.Date('2022-01-11')) %>%
   right_join(election, by = c('state' = 'state_name')) %>%
   select(date, state, incidence, party)
```

Second, test the equal variance assumption. 
```{r}
leveneTest(plt_states$incidence~plt_states$party, center = mean)
```
It shows that the variance of average incidence in different states grouped by vote result are significantly different. 

Next, do a T-test.

$H_0$ : The mean incidence in states with the two different election vote result are the same. 
Setting the confidence level as 0.95. 
```{r}
t.test(plt_states$incidence~plt_states$party, var.equal = F)
```
The p-value is far more less than 0.05. So the $H_0$ is rejected. I can conclude that there is difference in means of incidence between states who vote for Republican and states who vote for Democratic. 

In the following plot, we can also have an intuitive look that states who vote for Republican have higher incidence than states who vote for Democratic. 

```{r}
ggplot(states, aes(x = date, y = incidence, group = state)) + 
  geom_line(size = 0.1, alpha = 0.2) +
  geom_line(data = filter(states, state %in% gop), 
            aes(date, incidence, group = state, color = 'Republican'), 
            size = 0.3, alpha = 0.5) + 
  geom_line(data = filter(states, state %in% dem), 
            aes(date, incidence, group = state, color = 'Democratic'), 
            size = 0.3, alpha = 0.5) + 
  xlab('Jan 2020 - Oct 2021') + 
  ylab('Incidence') + 
  ggtitle('Figure 3: Incidence by States') + 
  scale_color_manual(values = c('Republican' = 'red', 'Democratic' = 'blue'))
```


# 5. Summary

This report analyze the trend of Covid-19 in the US both on country-level and state-level. With the plots, we can obviously see that there are three major peaks of cases from January 2020 to January 2022. The deaths trend basically accord with the cases trend, but with a little bit of lag. The states that have lower incidence before the outbreak in 2021's winter don't continue to have lower daily new cases in this new outbreak comparing with states that have higher incidence. 

The control of epidemic is affected by multiple factors. Given that the presidential election can reflect a political leaning of people in their region, the t-test on Covid-19 cases between states who vote for different parties is statistically significant. But the deeper reasons and the difference of public health policies between different states need further study. 

The data ends on `r max(us$date)`, on which the newly outbreak is still raging. This outbreak is different with previous outbreaks with exceedingly high infection speed, but the surge of deaths in this outbreak is relatively mild up to now. The further study on this outbreak and this variant virus is really important to get a better sense of the virus and to suppress the pandemic. 

